Leadership across cultures and contexts: an exploratory analysis

### 2020-03-20 Response to reviewers.Rmd

---
title: "Response to reviewers"
author: "Zach Garfield"
date: "3/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Response to editor and reviewer comments

### Editor comments

**Statistical reporting and further analysis: The GLMM should be reported in full somewhere. This could go in the supplemental materials. However, can we learn more from it? Do the random effects matter for the ethnographers? What about the national origin effects for the ethnographers? We recommend taking seriously the reflexive analytical techniques of putting the ethnographer in the analysis (see work by Robert Aunger working in the Congo Basin). Does the ethnographer matter? If so, how much? Further, and we realize there may not be sufficient power for this, but does the gender of the ethnographer predict ‘seeing’ female leaders in a population? This can (potentially) speak quantitatively to some of the qualitative discussion about limitations and biases in the approach.**

**Could further analyses be done to obtain more information on the nature of the variation within the data?**

**1.	The cluster analysis is nice, but it opens up the option to combine items within both qualities and functions to create a smaller number of higher-level categories or codes that, arguably, are better measured. The issue here is that some things that don’t appear frequently might rise in frequency when they are combined with other similar items. Suppose you use a higher-level category that contains 4 codes and you code the society as ‘1’ if any of the codes appears and zero otherwise. In our view, you need not go crazy with this approach, but look for any tightly bound clusters and consider pursuing those. If you are able to do this, the presentation could be more compact and easier to digest for the reader.**

**2.	Why not take either some of your original codes or some of these higher-level codes (from the above cluster analysis) and regress them on societal-level features related to subsistence type (foragers, pastoralist, etc. or simply forager vs. non-forager if that’s appropriate), levels of jurisdictional hierarchy (or other measures of societal complexity), population size, and potentially ecological variables. Your sample is not big, but you might be able to get some juice out of this. If interesting findings do emerge, you should consider adding or otherwise altering your visualizations.**

**3.	Be clearer about how you are accounting for research effort in each society. For example, are you controlling for the total number of words or pages for each population? This seems very important since we want to make sure that the variation observed isn’t due to some populations getting more or less total study effort or writing effort.**

**4.	Did you control for observation year? This might be too highly correlated with ethnographer identity, but worth checking**

**5.	Phylogenetic controls. While this is the probability sample, worries persist about non-independence (see R1). There are a variety of approaches to mitigate (and study!) this. Is there any clumping among related groups that cannot be explained by other variables? Do the patterns hold when you either control for linguistic distance or cluster in some way (continental fixed effects at a minimum)?**

**6.	Analyze the degree to which ‘researcher efforts’ (say measured by total words on that society and/or number of paragraphs on leadership) predict the appearance of a code. For example, what if the total ethnographic pages strongly predict the the appearance of “high status” as a code. Suppose the 10% of populations who don’t mention “high status” hardly wrote anything about leadership. On the other hand, if for other traits, there’s no relationship between researcher effort and appearance, then a case can be made that absence of evidence could very well be evidence of absence.**

**We recommend making liberal use of supplemental materials. Some of the analyses may not yield findings of sufficient interest to warrant placement in the main text, but some readers may want to have a look at them, so put them in the supplemental and briefly mention them in the main.**

**Returning to the WEIRD Problem: in your discussion, please circle back to the question of what research might be missing by focusing on leadership in Western or industrialized societies. Say more on how leaders in small-scale societies are the same and different—and perhaps you will be able to explain some of that variation with variables like jurisdictional hierarchy or subsistence type or mobility measures. As mentioned by R1, this likely involves returning to the GLOBE data and putting it together with your findings more directly.**

**Status vs. Leadership: R2 asks about status vs. leadership. Given recent debates about a “leadership psychology” as distinct and separable from status psychologies (like prestige and dominance, van Vugt), it might be worth expanding a bit on what these data say about the connection between status and leadership.**

**Expanding the dataset: the natural next step with this project is to use this small sample of hand coded categories to create an automated approach for all of the eHRAF, perhaps based on key words. The idea is to develop a set of keywords or other patterns that reliably predict when a paragraph should be coded in a particular way. The method can be developed and benchmarked using the dataset you’ve created and then simply applied in automated fashion to the entire eHRAF corpus. We offer this as a suggestion, but realize that it’s a lot to ask in this context.**

**Finally, the reviewers made a number of valuable suggestions, not enumerated above,  that should help you sharpen you presentation and discussion. Please attend to those.**

## Reviewer 1 comments: "Leadership across 59 traditional cultures: an exploratory analysis" Evolution & Human Behavior

**This manuscript presents a descriptive analysis of the HRAF to provide information about leadership in traditional societies. Overall, I enjoyed reading the paper and thought it was a useful resource. It provides a comprehensive analysis of leadership across a wide swath of traditional cultures. It reports on the types of functions leaders serve, the qualities possessed by leaders, and the costs and benefits of leadership experienced by both leaders and followers. The paper will likely be of interest to anyone working in the area of leadership, social hierarchy, or group behavior. I provide some comments below aimed at helping the authors increase the potential accessibility and impact of the work.**

**The paper was framed partially as providing an analysis of differences between traditional societies and WEIRD societies but, unless I missed it, there was not much analysis along these lines. Neither the background section nor the Discussion section focuses much on contrasting WEIRD and traditional societies. The paper appears to miss out on a useful opportunity to provide some commentary on this issue. Also, I realize the paper is decidedly exploratory/descriptive, but did the authors have any hypotheses about differences between WEIRD and traditional societies?**

**The paper suggests that one of the most important qualities and benefits of being a leader involves being of "high status." That's a little dissatisfying. What's the difference between being high status and being a leader? I realize the authors provided brief operational definitions of the two constructions, but might they simply be two ways of describing the same thing (having a high degree of influence)? Perhaps the authors could do just a bit more to differentiate them.**

**Some of the benefits of leadership also appear as costs. Greater reflection and explanation on this aspect of the data seems warranted. I understand that leaders can enjoy certain benefits (access to resources) but then if they lose their status as leader they might lose those benefits (if someone falls from grace, they no longer get a larger piece of the pie). Do the findings reflect anything more complex than that? Their discussion of benefits of costs was brief and a little disappointing. The paper might benefit from enhancing that discussion.**

**Some readers might appreciate a bit more detail about how the HRAF database was created, including how the ethnographic records were identified and selected and whom and when they are from.**

**I don't work with ethnographic data, so perhaps this question is naïve, but: What conclusions can we draw about the actual frequency of leadership functions and traits from the frequency with which they are mentioned in the ethnographic record? There is an implicit assumption that the two are closely aligned. Has such an alignment been confirmed? Is that even testable?**

**What does it mean for someone to be a leader at the "supracommunity" level? I'm not familiar with the term, so had difficulty contextualizing the relevant findings. And how exactly do the authors define "community" versus "supracommunity"?**

## Reviewer 2 comments

**The authors present results of a Herculean task of mining the ethnographic record to estimate frequency of and covariation between functions, qualities, and costs/benefits of leadership. While the study itself is exploratory, it will be very useful to future work testing models of the evolution of leadership and social inequality in general. I have a number of suggestions and comments that may improve the paper, none highly critical:**

**The abstract could be tightened up in terms of description of particular costs and benefits to leaders and followers. And a brief justification in the abstract for why the study compares costs and benefits would be useful- e.g. understanding evolution of leader or follower strategies requires understanding the (fitness-relevant) costs and benefits to them.**

**Lines 88-89 suggest that the paper's ethnographic analysis will include leadership at all levels of social organization. But does the theory and analysis reported in the paper really extend to all levels of social organization- such as child play groups or households? It seems the focus is still community-level cooperation among adults. Perhaps clarify.**

**In the "Functions of leadership" section, I would expand on your claim that leadership emerges when there is demand for a "managerial presence". You might describe collective action or coordination problems here. That might better motivate the subsequent discussion of different functions of leadership in small-scale societies, as well as discussion of the conditions (e.g. demographic change, wealth access) that increase institutionalization or authority of leaders. Glowacki and von Rueden (2015) suggest that inter-family conflict resolution is where leader institutionalization often first arises, and Roscoe (2009) makes a similar argument concerning leadership in highland New Guinea, where the institutionalized display of material wealth organized by Big Men is rewarded more than ad hoc leadership during war.**

**The Wofford study (lines 214-233) and GLOBE study reminds me of the vast array of descriptive models in leadership/managerial studies- transactional vs. transformational leadership and their sub-components being one of the most popular. So what should a reader make of this literature, with respect to evolution of leadership in small-scale societies? Do bureaucracies, formal legal systems, formal education-based skills, and large inequalities in remuneration limit universality of descriptive models from leadership/managerial studies? I'd be more explicit on this in the introduction. For the discussion section: how does your own descriptive modeling of leadership in traditional societies agree or not with Wofford or other models from large-scale societies?**

**There is little mention of gender in the Qualities of leadership section, even though this might be first attribute an alien might recognize as covarying with community-level leadership in human societies. Even in the more egalitarian societies, men often wield more overt political influence during camp or band-wide political discussions. Without getting into arguments about patriarchy and its origins, it might at least be useful to cite your own work on Chabu and work on the Tsimane (von Rueden et al. 2018) that finds similar traits predict leadership for men and women, implying its not gender per se that gives men a leadership advantage in politics.**

**How often are shamans also the most influential political figure (given your suggestion in line 317)? My sense from various ethnographies is that this was not the majority of cases (though often the case), whether in more egalitarian societies or chiefdoms. I could be wrong.**

**Even though this paper is more exploratory, in the last section before the methods, can you venture any predictions about leader function-leader quality covariation? Or about how reported costs and benefits of leadership and followership may vary depending on subsistence type?**

**Many of the variables coded as qualities of leaders may also fall under leader costs and benefits, e.g. polygyny, physical health, wealthy. Were variables like these analyzed in both cases?**

**Can you control for phylogeny (e.g. linguistic affinity). See Minocher et al. 2019 in Evolution and Human Behavior. It may be that covariation you detect between qualities and leader functions is driven by shared cultural history more than direct causal relationship between qualities and leader functions.**

**Can you report text records pertaining to male vs. female leadership by subsistence type (lines 966-982)?**

**While your analysis strategy was not designed to test particular predictions, I'd be very interested to see you report how the authority variable covaries with different leader functions in different subsistence types, which has relevance to institutionalization of leadership and evolution of social inequality.**

**Give more explanation of prestige theory and computational services theory (lines 1279-1280), including what different predictions they make.**

**Followers in many cases may be making best of limited options, particularly in societies with steeper hierarchy. So I'm not sure benefits of followership must outweigh the costs (1414-1415) in same way as for leadership.**

### 2020-04-02 Response to reviewers.Rmd

**Finally, the reviewers made a number of valuable suggestions, not enumerated above,  that should help you sharpen you presentation and discussion. Please attend to those.**

## Reviewer 1 comments

**This manuscript presents a descriptive analysis of the HRAF to provide information about leadership in traditional societies. Overall, I enjoyed reading the paper and thought it was a useful resource. It provides a comprehensive analysis of leadership across a wide swath of traditional cultures. It reports on the types of functions leaders serve, the qualities possessed by leaders, and the costs and benefits of leadership experienced by both leaders and followers. The paper will likely be of interest to anyone working in the area of leadership, social hierarchy, or group behavior. I provide some comments below aimed at helping the authors increase the potential accessibility and impact of the work.**

We greatly appreciate the reviewers detailed comments and very useful suggestions.

**The paper was framed partially as providing an analysis of differences between traditional societies and WEIRD societies but, unless I missed it, there was not much analysis along these lines. Neither the background section nor the Discussion section focuses much on contrasting WEIRD and traditional societies. The paper appears to miss out on a useful opportunity to provide some commentary on this issue. Also, I realize the paper is decidedly exploratory/descriptive, but did the authors have any hypotheses about differences between WEIRD and traditional societies?** 
We agree our initial submisson did not sufficently accomplish this goal. We have now added summarizing, concluding statements on the qualities and functions of leaders in WEIRD and non-WEIRD settings in the introduction.

NEED TO ADD TO DISCUSSION & ANALYSES?

@@

**The paper suggests that one of the most important qualities and benefits of being a leader involves being of "high status." That's a little dissatisfying. What's the difference between being high status and being a leader? I realize the authors provided brief operational definitions of the two constructions, but might they simply be two ways of describing the same thing (having a high degree of influence)? Perhaps the authors could do just a bit more to differentiate them.**

<!-- IDEA: Perform some textual analysis on teh 10% of text records in which there was not evidence leaders were high status. How does these records compare RE gender, group structure, or other qualities (maybe), or the most frequent functions, i.e., try to charachterize non-high status leaders.  -->

@@

**Some of the benefits of leadership also appear as costs. Greater reflection and explanation on this aspect of the data seems warranted. I understand that leaders can enjoy certain benefits (access to resources) but then if they lose their status as leader they might lose those benefits (if someone falls from grace, they no longer get a larger piece of the pie). Do the findings reflect anything more complex than that? Their discussion of benefits of costs was brief and a little disappointing. The paper might benefit from enhancing that discussion.**

<!-- IDEA: Represent the benefits and costs of leadership within group type. Would need to standardize the estimates, such our culture scores in first study, but it could be group structure benefit score, etc. Then we can roughly compare how costs and benefits comapre across group types. -->

@@

**Some readers might appreciate a bit more detail about how the HRAF database was created, including how the ethnographic records were identified and selected and whom and when they are from.**

We added a few additional details in the methods section on the eHRAF and ethngraphic sample and now more clearly direct readers to the original pubication from which the database was built, which includes additional details.

We also include additional details in the results and supplimentary information on the corpus of ethnographic documents which provide the sample of ethnographic paragraphs. NEED TO DO.

@@

**I don't work with ethnographic data, so perhaps this question is naïve, but: What conclusions can we draw about the actual frequency of leadership functions and traits from the frequency with which they are mentioned in the ethnographic record? There is an implicit assumption that the two are closely aligned. Has such an alignment been confirmed? Is that even testable?**

The reviewer brings up an important point. For one, this is partly our motivation for representing the percentage of cultures which provide at least one text record supporting each of the operationalized variables. Despite this low threshold for acceptign supporting evidence, the ethnogrtaphic record (sampled here) does not provide support for every variable from every culture. At this culture-level, we feel more confident our data provide a estiamte of cross-cultural frequency, albeit an inherently noisy estimation. Also, the entire field of ethnology (the branch of anthropology focused on comparing charachteristics of populations across the globe), is based on developing comparable and often quantitaive measures from primary ethnogrpahic information. For example, the Ethnographic Atlas, which provides hundreds of coded quantitaive and qualitive variables on thousands of cultures and the Standard Cross-Cultural Sample, which provides thousands of coded quantitative and qualitative variables on 186 cultures, were developed in essentially the same way we have developed our operatiopnalized variables here. Developing and contrasting coded measures from primary ethnographic documents is not without limitations, but does represent (perhaps) the only method to systematically and empirically compare measures across a large sample of cultures/populations.

At the text-record level drawing strong conclusions about the actual frequency of traits is slightly more problematic. However, fundamentally these results provide a systematic view of the ethnographic record. From these results field researchers and theoreticians can draw inference from a more comprehensive view of human culturla diversity rather than haphazardly self-selecting a few ethngoraphic cases studies.

We now discuss these limitations and points more clearly in the limitations setion. NEED TO DO.

@@

**What does it mean for someone to be a leader at the "supracommunity" level? I'm not familiar with the term, so had difficulty contextualizing the relevant findings. And how exactly do the authors define "community" versus "supracommunity"?**

In the methods section we now include a description of all levels of our *group structure" coding to clearly explain this coding scheme. We define supracommunity leaders when the "group" involves leadership beyond the community level and across multiple residential communities, such as regional political leaders, chiefs in complex chiefdoms, kings, and state-level leaders.

### 2020-04-02 leadership_across_cultures_contexts.Rmd
Starting to work on response to reviewers.
Adding cost v benefits code.

---
title: "Leadership across 59 traditional cultures: an exploratory analysis"
author: "Zachary H. Garfield, Kristen L. Syme, and Edward H. Hagen"
date: "4/25/2019"
output: 
  bookdown::word_document2
bibliography: Leader2.bib
csl: evolution-and-human-behavior.csl
---

@@

We first briefly review theoretical, ethnographic, and empirical perspectives on the qualities of leaders, the functions they serve, and the costs and benefits of leadership across human societies, including WEIRD and non-WEIRD populations. We then outline the aims and methods of the current cross-cultural study, before presenting results and interpretations. Through novel exploratory analyses encompassing the full range of cultural diversity and social contexts, we aim to advance leadership studies across disciplines and provide researchers with a foundation for more detailed approaches to human leadership across cultures and contexts.

@@

<!-- NEED CONCLUSION PARAGRAPH/STATEMENT.  -->

@@

Across culturally and economically diverse populations there are some broad commonalities in the qualities of community leaders. When strong community leaders emerge, they are most typically high status males who embody the ideal qualities most respected by the group and necessary for successful leadership in the domains in which they have influence [@henrich_evolution_2001; @van_vugt_evolutionary_2006; @collier_politics_1981; @low_men_1992; @roscoe_social_2019]. Elders, respected individuals, and highly intelligent or knowledgeable individuals tend to have greater influence over community decisions across cultures, including among the most egalitarian hunter-gatherers, stratified chiefdoms, and contemporary state-level organizations [@silverman_how_1978; @van_vugt_cognitive_2007; @tooby_cognitive_2006; @garfield_review_2019; @bass_handbook_1990]. Drawing on cross-national data through the GLOBE project from over 15,000 middle managers across 60 countries, @den_hartog_culture_1999 conclude that elements of charismatic and transformational leadership -- leadership styles when leaders express an inclusive vision of future progress, lead through inspiration, and develop initiatives in collaboration -- are universally endorsed across nations and are perceived to be associated with effective leadership. In particular, inducing motivation, having foresight, being encouraging and communicative, trustworthiness, and building self-confidence among followers are suggested to be universal attributes of outstanding leadership in managerial contexts.

@@

<!-- NEED CONCLUSION PARAGRAPH/STATEMENT. -->

@@

We evaluate the cross-cultural frequency in the functions and qualities of leaders and the costs and benefits of leadership for both leaders and followers using an ethnographic database previously built and published (doi.org/10.5281/zenodo.2541999), from which @garfield_evolutionary_2019 "tested" three theoretical models against the ethnographic record. Whereas @garfield_evolutionary_2019, documented supporting evidence for measures of leadership in reference to influential theoretical models, our goals here are to let the ethnographic record speak for itself in understanding variation in the dimensions of leadership we are interested in, i.e., our methods and analyses are exploratory and not strictly theoretically driven.

Here we draw on the same database of 1212 ethnographic paragraphs (termed text records) extracted from the 60-culture Probability Sample Files (PSF) of the electronic Human Relations Area Files (eHRAF) database [@naroll_proposed_1967]. The *leadershipdata* database was designed to capture a wide sample of the ethnographic record of leadership. For the current study, two authors coded each text record for presence of `r length(quality_vars)` distinct qualities of leaders, `r length(function_vars)` distinct functions leaders serve, and `r length(leader_benefit_vars)` benefits and `r length(leader_cost_vars)` costs that leaders or followers may receive.

The operationalized qualities and functions (See Table \@ref(tab:qualfuncvars)) were developed from the ethnographic text records in the database. First ZG read a subset of text records and developed initial variable lists and operationalizations of the qualities and functions in those text records. Then, KS coded all 1212 text records (in a random order) for the operationalized qualities, functions, cost, and benefits, while also creating new variables and operationalizations for additional qualities and functions encountered. After coding half of the random sample KS and ZG revised the operational definitions to produce the final list of variables in Table \@ref(tab:qualfuncvars). ZG then followed the same process of coding the entire sample of 1212 text records for the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` operationalized variables. ZG and KS then discussed every instance of disagreement between their respective coding to produce a consensus matrix, which is used in the analyses here and added to the published data repository (a data frame named *leader_text2*).

@@

```{r cluster, fig.cap="Cluster analysis of leader quality variables (left) and leader function variables (right). Distances are $1 - cor$. Ward agglomeration method. AU p-values (red) computed with 10,000 bootstrap samples using the pvclust package [@pvclust2015]."}
# plot(m_pvclust_qual)
# pvrect(m_pvclust_qual, alpha = 0.9)
knitr::include_graphics("Figures/m_pvclust.both.svg")
```

>>>>

---
title: "Leadership across cultures and contexts: an exploratory analysis"
author: "Zachary H. Garfield, Kristen L. Syme, and Edward H. Hagen"
date: "4/25/2019"
output: 
  bookdown::html_document2
bibliography: Leader2.bib
csl: evolution-and-human-behavior.csl
---

@@

We first briefly review theoretical, ethnographic, and empirical perspectives on the qualities of leaders, the functions they serve, and the costs and benefits of leadership across human societies, including WEIRD and non-WEIRD populations [i.e., @henrich_weirdest_2010]. We then outline the aims and methods of the current cross-cultural study, before presenting results and interpretations. Through novel exploratory analyses encompassing the full range of cultural diversity and social contexts, we aim to advance leadership studies across disciplines and provide researchers with a foundation for more detailed approaches to human leadership across cultures and contexts.

@@

<!-- ADDED CONCLUSION PARAGRAPH/STATEMENT.  -->
In contrasting the functions leaders serve in smaller-scale, non-industrial, and more traditional societies (i.e., non-WEIRD), from larger-scale, (post)industrial contexts (i.e., WEIRD), community leaders in non-WEIRD settings are more likely to have domain-specific functions based on role-specialization [see @smith_leadership_2016]. The hunt leader does not necessarily have authority over military raids and the lineage head does not necessarily have authority over ritual ceremonies. In WEIRD settings however, community, organizational, and political leaders often have decision-making authority across multiple domains. In the United States, despite a designed seperation of power across branches of government, presidents, governors, and mayors oversee governmental bodies and have substantial influence in economics, social and legal policies, and international and inter-regional relations. Community leadership functions in increasingly WEIRD contexts will likely be increasingly focused on economics and inter-group relations, whereas in less WEIRD contexts community leaders will primarily function to resolve interpersonal conflicts and manage inter-group social dynamics. 
<!-- Could add ethnographic examples for SSS and LSS  -->

@@

Across culturally and economically diverse populations there are some broad commonalities in the qualities of community leaders. When strong community leaders emerge, they are most typically high status males who embody the ideal qualities most respected by the group and necessary for successful leadership in the domains in which they have influence [@henrich_evolution_2001; @van_vugt_evolutionary_2006; @collier_politics_1981; @low_men_1992; @roscoe_social_2019]. Elders, respected individuals, and highly intelligent or knowledgeable individuals tend to have greater influence over community decisions across cultures, including among the most egalitarian hunter-gatherers, stratified chiefdoms, and contemporary state-level organizations [@silverman_how_1978; @van_vugt_cognitive_2007; @tooby_cognitive_2006; @garfield_review_2019; @bass_handbook_1990]. Drawing on cross-national data through the GLOBE project from over 15,000 middle managers across 60 countries, @den_hartog_culture_1999 conclude that elements of charismatic and transformational leadership -- leadership styles when leaders express an inclusive vision of future progress, lead through inspiration, and develop initiatives in collaboration -- are universally endorsed across nations and are qualities perceived to be associated with effective leadership. In particular, inducing motivation, having foresight, being encouraging and communicative, trustworthiness, and building self-confidence among followers are suggested to be universal attributes of outstanding leadership in managerial contexts.

@@

<!-- NEED CONCLUSION PARAGRAPH/STATEMENT. -->
Much of the theoretical leadership literature across disciplines suggests relative uniformity in many qualities of community leaders. Intelligence, charisma, effective decision-making capabilities, high social status and strong social networks are likely to be universal qualities of political and community leaders. In more socially stratified, politically complex, and in increasingly WEIRD contexts, heredity, ascribed social statuses, and affiliations are more likely to be critical qualities in achieving and/or maintaining leadership positions. We also envision a curvilinear relationship between leadership status and supernatural abilities and qualities. Community leaders among small-scale hunter-gatherers may be seen as experts in the use of ritual and supernatural "tools" though they typically do not have a monopoly in their use. In chiefdoms and and in early prehistoric state societies political leaders are often seen as divine and have special supernatural qualities commoners do not. In contemporary WEIRD societies there is a trend of secularization among political leaders, generally, although religious and spiritual leaders do have substantial influence over social and moral norms.

@@

We estimate the cross-cultural frequency in the functions and qualities of leaders and the costs and benefits of leadership for both leaders and followers using an ethnographic database previously built and published (doi.org/10.5281/zenodo.2541999), from which @garfield_evolutionary_2019 "tested" three theoretical models against the ethnographic record. The electronic Human Relations Area Files (eHRAF) is a digital collection of ethnographies, subject-coded at teh paragraph level following a the Outline of Cultural Materials (OCM) coding scheme. The eHRAF is a subset of the larger HRAF and is continually growing in scope.

Whereas @garfield_evolutionary_2019, documented supporting evidence for measures of leadership in reference to influential theoretical models, our goals here are to let the ethnographic record "speak for itself" in understanding variation in the dimensions of leadership we are interested in, i.e., our methods and analyses are exploratory and not strictly theoretically driven. We draw on the same database of 1212 ethnographic paragraphs (termed text records) extracted from the 60-culture Probability Sample Files (PSF) of the eHRAF database [@naroll_proposed_1967]. The *leadershipdata* database was designed to capture a wide sample of the ethnographic record of leadership across social contexts and was built using an advanced search including 16 OCM codes related to leadership seven key words describing leaders or followers. For the current study, two authors coded each text record for presence of `r length(quality_vars)` distinct qualities of leaders, `r length(function_vars)` distinct functions leaders serve, and `r length(leader_benefit_vars)` benefits and `r length(leader_cost_vars)` costs that leaders or followers may receive. See @garfield_evolutionary_2019 for additional details on the ethnographic sample and methodology.

The operationalized qualities and functions (See Table \@ref(tab:qualfuncvars)) were developed from the ethnographic text records in the database. First ZG read a subset of text records and developed initial variable lists and operationalizations of the qualities and functions in those text records. Then, KS coded all 1212 text records (in a random order) for the operationalized qualities, functions, cost, and benefits, while also creating new variables and operationalizations for additional qualities and functions encountered. After coding half of the random sample KS and ZG revised the operational definitions to produce the final list of variables in Table \@ref(tab:qualfuncvars). ZG then followed the same process of coding the entire sample of 1212 text records for the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` operationalized variables, as well as a measure of group structure coded as: *residential subgroup*, which includes informal groups of co-residents, age-based groups, performance groups; *kin group*, which includes lineages, phraties, and other kin-based organizations; *economic group*, which includes subsistence groups, market groups, and other groups with primarily economic goals; *military group*, which includes all groups related to inter-group conflcit; *religious group*, which includes groups collected for spiritual or supernatural purpose; *political group (community)*, which includes political groups at the level fo the community, i.e., the political leader can potentially directly interact with most community members such as most village headmen, or *political group (supracommunity)*, which includes political groups that encompass multiple residential communities, such as complex chiefdoms, regional political leaders, and kings or state-level leaders. ZG and KS then discussed every instance of disagreement between their respective coding to produce a consensus matrix, which is used in the analyses here and added to the published data repository (a data frame named *leader_text2*).

@@

```{r cluster, fig.cap="Cluster analysis of leader quality variables (left) and leader function variables (right). Distances are $1 - cor$. Ward agglomeration method. AU p-values (red) computed with 10,000 bootstrap samples using the pvclust package [@pvclust2015]."}
# plot(m_pvclust_qual)
# pvrect(m_pvclust_qual, alpha = 0.9)
knitr::include_graphics("Figures/m_pvclust.both.png")
```
